---
title: "Paired difference in protein aggregation versus protein age inferred from regression analysis "
author: "Ben Wilson"
output: html_document
---
  
Load mySQL table info for genes
```{r}
gene.data <- read.table("InputFinal/EnsemblV73_MouseGenes.tab", sep="\t",header=T) 
protein.data <- read.table("InputFinal/EnsemblV73_MouseProteins.tab", sep="\t",header=T)

APR.data <- merge(gene.data, protein.data, by.x="MouseGenesPrimaryKey", by.y="MouseGenesForeignKey")
remove(gene.data)
remove(protein.data)

scrambled.data <- read.table("InputFinal/FILVMW_2018ScrambledControls.tab",sep="\t",header=TRUE)

```

```{r}
APR.data[which(APR.data$EnsemlGeneUID=="ENSMUSG00000067543"),]$PS21_GeneFamilyNumber <- APR.data[which(APR.data$EnsemlGeneUID=="ENSMUSG00000052262"),]$PS21_GeneFamilyNumber #Fix gene family for two pahari genes

paralogs.list <- c("ENSMUSG00000073631","ENSMUSG00000079800","ENSMUSG00000094728","ENSMUSG00000095240","ENSMUSG00000095366","ENSMUSG00000095450","ENSMUSG00000095787","ENSMUSG00000096100","ENSMUSG00000096850")
APR.data <- APR.data[which(APR.data$PS21_GeneFamilyPhylostratum != "0" & APR.data$PS21_GeneFamilyPhylostratum != "21" & APR.data$PS21_ExcludedGeneBinary != "1" ),] #exclude unclassifiable genes, false-positive rich ORFans
APR.data <- APR.data[which(!(APR.data$EnsemblGeneUID %in% paralogs.list)),]#remove genes with known paralogy to other genes


```


Factorize gene family number and phylostratum
```{r}
APR.data$PS21_GeneFamilyNumber <- as.factor(APR.data$PS21_GeneFamilyNumber) #factorize gene family ID to prevent numeric interpretation by R
APR.data$PS21_GeneFamilyPhylostratum <- as.factor(APR.data$PS21_GeneFamilyPhylostratum) #factorize variable to ensure that it is not interpreted as numeric variable
APR.data$PS21_GeneFamilyPhylostratum <- with(APR.data, factor(PS21_GeneFamilyPhylostratum, levels = c(sort(unique(APR.data$PS21_GeneFamilyPhylostratum))))) # set levels of categorical phylostratum variable
APR.data$PS21_ExcludedGeneBinary <- as.factor(APR.data$PS21_ExcludedGeneBinary) #factorize binary category for gene families with ambiguous age


```


Split data frames into categories
```{r}
genes <- APR.data[which(APR.data$GeneOrControlDesignation=="CodingGene"),c("EnsemblGeneUID","PS21_GeneFamilyPhylostratum","PS21_GeneFamilyNumber","TangoPentamerAAInAPRsDensity","WaltzPentamerAAInAPRsDensity")] #split out genes
scrambled <- APR.data[which(APR.data$GeneOrControlDesignation=="ScrambledByAminoAcidControl"), c("EnsemblGeneUID","TangoPentamerAAInAPRsDensity","WaltzPentamerAAInAPRsDensity")] #...scrambled controls,
dispersion <- APR.data[which(APR.data$GeneOrControlDesignation=="ConsistentHydrophobicDispersionControl"), c("EnsemblGeneUID","TangoPentamerAAInAPRsDensity","WaltzPentamerAAInAPRsDensity")] #...and dispersion controls

#USE FOR TANGO
scrambled.mean <- aggregate(scrambled.data$TangoPentamerAAInAPRsDensity, by = list(scrambled.data$EnsemblGeneUID), FUN = mean, data=scrambled.data)
colnames(scrambled.mean) <- c("EnsemblGeneUID","ScrambledMeanAPR")

#USE FOR WALTZ
# scrambled.mean <- aggregate(scrambled.data$WaltzPentamerAAInAPRsDensity, by = list(scrambled.data$EnsemblGeneUID), FUN = mean, data=scrambled.data)
# colnames(scrambled.mean) <- c("EnsemblGeneUID","ScrambledMeanAPR")


remove(APR.data) #clear bulk data frame

#Use these for TANGO
colnames(genes)[4] <- "GeneAPR" #make ISD column names unique so that they can be merged by EnsemblGeneUID
colnames(scrambled)[2] <- "ScrambledAPR" #same here
colnames(dispersion)[2] <- "DispersionAPR" #same here

#Use these for Waltz
# colnames(genes)[5] <- "GeneAPR" #make ISD column names unique so that they can be merged by EnsemblGeneUID
# colnames(scrambled)[3] <- "ScrambledAPR" #same here
# colnames(dispersion)[3] <- "DispersionAPR" #same here

```

Merge genes and paired controls by EnsemblGeneID, combine into single data frame for calculating pairwise difference
```{r}
paired.APR.df.1 <- merge(genes, scrambled, by = "EnsemblGeneUID", all.x = TRUE) #merge genes and scrambled first
paired.APR.df.2 <- merge(paired.APR.df.1, dispersion, by = "EnsemblGeneUID", all.x = TRUE) #merge in dispersion
APR.df <- merge(paired.APR.df.2, scrambled.mean, by = "EnsemblGeneUID", all.x = TRUE) #merge the previous with dispersion controls (R merge only allows pairwise mergers)
remove(genes) 
remove(scrambled) 
remove(dispersion)
remove(paired.APR.df.1)
remove(paired.APR.df.2)
```

Construct comparison for the difference between means of scrambled and unscrambled genes and dispersion controls
```{r}
#Construct a data frame for the difference in means between genes and scrambled genes
APR.df$DifferenceGeneVsScrambledAPR <- APR.df$GeneAPR - APR.df$ScrambledAPR
APR.df$DifferenceGenesVsDispersionAPR <- APR.df$GeneAPR - APR.df$DispersionAPR
APR.df$DifferenceGenesVsScrambledMean <- APR.df$GeneAPR - APR.df$ScrambledMeanAPR
APR.df$DifferenceOfDifferences <- APR.df$DifferenceGenesVsScrambledMean - APR.df$DifferenceGenesVsDispersionAPR
```

Generate family of linear models for each phylostratum, test significance of gene vs. randomized control controlling for gene family variance
```{r}
library(nlme)

lm.APR.genes.scrambled <- lme(DifferenceGeneVsScrambledAPR ~ PS21_GeneFamilyPhylostratum, random = ~ 1|PS21_GeneFamilyNumber, data = APR.df) #linear mixed-effects model for scrambled difference with gene families as independent sources of error
lm.APR.genes.dispersion <- lme(DifferenceGenesVsDispersionAPR ~ PS21_GeneFamilyPhylostratum, random = ~ 1|PS21_GeneFamilyNumber, data =APR.df) #linear mixed-effects model for GC difference with gene families as independent sources of error
lm.APR.scrambled.dispersion <- lme(DifferenceOfDifferences ~ PS21_GeneFamilyPhylostratum, random = ~ 1|PS21_GeneFamilyNumber, data =APR.df) #linear mixed-effects model for GC difference with gene families as independent sources of error

lm.APR.genes.scrambled.mean <- lme(DifferenceGenesVsScrambledMean ~ PS21_GeneFamilyPhylostratum, random = ~ 1|PS21_GeneFamilyNumber, data =APR.df) 


p.value.levels <- c(summary(lm.APR.scrambled.dispersion)$tTable[,5] <0.01) + c(summary(lm.APR.scrambled.dispersion)$tTable[,5] <0.001) + c(summary(lm.APR.scrambled.dispersion)$tTable[,5] <0.0001)
p.value.vector <- strrep("*",p.value.levels)
significant.ps <- unname(which(p.value.levels != 0))
insignificant.ps <- setdiff(1:20,significant.ps)



```


Plot means from each model with confidence intervals to test if they are significantly different
```{r}
# Put model estimates into temporary data frames
last.index.scrambled <- length(fixed.effects(lm.APR.genes.scrambled)) #get number of phylostrata that have been fit
genes.scrambled.df <- data.frame(Phylostratum = names(fixed.effects(lm.APR.genes.scrambled)[1:last.index.scrambled]), Intercept = unname(fixed.effects(lm.APR.genes.scrambled)[1:last.index.scrambled]), SE = unname(summary(lm.APR.genes.scrambled)$tTable[,2][1:last.index.scrambled]), Model = "Genes relative to scrambled control") #construct data frame to hold linear model results
genes.scrambled.df$Intercept[2:last.index.scrambled] <- genes.scrambled.df$Intercept[2:last.index.scrambled] + genes.scrambled.df$Intercept[1] #increment means by first intercept to correct for R convention of relative means


last.index.scrambled.mean <- length(fixed.effects(lm.APR.genes.scrambled.mean)) #get number of phylostrata that have been fit
genes.scrambled.mean.df <- data.frame(Phylostratum = names(fixed.effects(lm.APR.genes.scrambled.mean)[1:last.index.scrambled.mean]), Intercept = unname(fixed.effects(lm.APR.genes.scrambled.mean)[1:last.index.scrambled.mean]), SE = unname(summary(lm.APR.genes.scrambled.mean)$tTable[,2][1:last.index.scrambled.mean]), Model = "Genes relative to scrambled controls") #construct data frame to hold linear model results
genes.scrambled.mean.df$Intercept[2:last.index.scrambled.mean] <- genes.scrambled.mean.df$Intercept[2:last.index.scrambled.mean] + genes.scrambled.mean.df$Intercept[1] #increment means by first intercept to correct for R convention of relative means


last.index.dispersion <- length(fixed.effects(lm.APR.genes.dispersion)) #same for dispersion controls...
genes.dispersion.df <- data.frame(Phylostratum = names(fixef(lm.APR.genes.dispersion)[1:last.index.dispersion]), Intercept = unname(fixef(lm.APR.genes.dispersion)[1:last.index.dispersion]), SE = unname(summary(lm.APR.genes.dispersion)$tTable[,2][1:last.index.dispersion]), Model = "Genes relative to hydrophobic clustering controls") #...
genes.dispersion.df$Intercept[2:last.index.dispersion] <- genes.dispersion.df$Intercept[2:last.index.dispersion] + genes.dispersion.df$Intercept[1] #...


last.index.scrambled.dispersion <- length(fixed.effects(lm.APR.scrambled.dispersion)) #same for dispersion controls...
scrambled.dispersion.df <- data.frame(Phylostratum = names(fixef(lm.APR.scrambled.dispersion)[1:last.index.scrambled.dispersion]), Intercept = unname(fixef(lm.APR.scrambled.dispersion)[1:last.index.scrambled.dispersion]), SE = unname(summary(lm.APR.scrambled.dispersion)$tTable[,2][1:last.index.scrambled.dispersion]), Pvalue = summary(lm.APR.scrambled.dispersion)$tTable[,5][1:last.index.scrambled.dispersion], Model = "Paired difference of differences" , Pstars = p.value.vector) #...
scrambled.dispersion.df$Intercept[2:last.index.scrambled.dispersion] <- scrambled.dispersion.df$Intercept[2:last.index.scrambled.dispersion] + scrambled.dispersion.df$Intercept[1] #...




percent.change.in.difference <- scrambled.dispersion.df$Intercept/genes.scrambled.mean.df$Intercept * 100
percent.change.in.ci <- genes.scrambled.mean.df$SE/genes.scrambled.df$SE * 100
percent.reduction.in.ci <- 100.00 - percent.change.in.ci


genes.scrambled.df$Phylostratum <- replace(as.character(genes.scrambled.df$Phylostratum),1:last.index.scrambled,c(1:last.index.scrambled)) #replace linear model text with factorized phylostrata
genes.scrambled.df$Phylostratum <- with(genes.scrambled.df, factor(Phylostratum, levels = c(1:last.index.scrambled))) #set levels of factorized PS


genes.scrambled.mean.df$Phylostratum <- replace(as.character(genes.scrambled.mean.df$Phylostratum),1:last.index.scrambled.mean,c(1:last.index.scrambled.mean)) #replace linear model text with factorized phylostrata
genes.scrambled.mean.df$Phylostratum <- with(genes.scrambled.mean.df, factor(Phylostratum, levels = c(1:last.index.scrambled.mean))) #set levels of factorized PS



genes.dispersion.df$Phylostratum <- replace(as.character(genes.dispersion.df$Phylostratum),1:last.index.dispersion,c(1:last.index.dispersion)) #same for dispersion data frame
genes.dispersion.df$Phylostratum <- with(genes.dispersion.df, factor(Phylostratum, levels = c(1:last.index.dispersion))) # set levels of factorized PS



scrambled.dispersion.df$Phylostratum <- replace(as.character(scrambled.dispersion.df$Phylostratum),1:last.index.scrambled.dispersion,c(1:last.index.scrambled.dispersion)) #same for dispersion data frame
scrambled.dispersion.df$Phylostratum <- with(scrambled.dispersion.df, factor(Phylostratum, levels = c(1:last.index.scrambled.dispersion))) # set levels of factorized PS





combined.differences.df <- rbind(genes.scrambled.df, genes.scrambled.mean.df, genes.dispersion.df) #combine data frames into single result data frame

```

Create colorblind friendly color scheme and reverse transform function
```{r}

#Create colorblind friendly palette
colorblindPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
colBlack <- colorblindPalette[1]
colOrange <- colorblindPalette[2]
colSkyBlue <- colorblindPalette[3]
colGreen <- colorblindPalette[4]
colYellow <- colorblindPalette[5]
colDarkBlue <- colorblindPalette[6]
colVermillion <- colorblindPalette[7]
colPink <- colorblindPalette[8]

#Create function for reverse box-cox transform to retransform data onto readable axis (note that this preserves relative comparison of certainty between PS but does not preserve absolute measure of SEM)
bc.reverse.transform <- function(x,l1,l2){
  x.reverse.transform <- (x*l1+1)^(1/l1)-l2
  return(x.reverse.transform)
}

```

Make plots to visualize differences in paired aggregation
```{r}
# Plots for each categorization
library("ggplot2")
library("grid")

interval.times <- read.table("split_interval_times.txt")
axis.breaks <- -log10(interval.times$V1)

split.labels<-sprintf("%.2E", interval.times$V1)

label.vector <- c("Cellular org.","Eukaryota","Opisthokonta","Holozoa","Metazoa","","","","","","Vertebrata","Euteleostomi","Tetrapoda","Amniota","Mammalia","Eutheria","Boreoeutheria\n","Euarchontoglires","Rodentia","Mus")
split.times <- as.character(round(interval.times$V1/1000000.0))



combined.differences.df$Times <-c(rep(axis.breaks,3))
percent.labels <- paste(as.character(round(percent.change.in.difference)),"%",sep="")
combined.labels <- paste(percent.labels,p.value.vector,sep="")

color.vector <- rep("black",20)
color.vector[significant.ps] <- "red"

top.positions <- combined.differences.df$Intercept + 2.5*combined.differences.df$SE

bottom.positions <- combined.differences.df$Intercept - 2*combined.differences.df$SE

plot.APR.difference <- ggplot(combined.differences.df[which(combined.differences.df$Model!="Genes relative to scrambled control"),], aes(colour = Model, fill= Model, shape = Model)) #Create ggplot object
plot.APR.difference <- plot.APR.difference + 
  geom_hline(yintercept = 0, linetype = "dashed", colour = "black") + #add horizontal line for null expectation at 0
  
  geom_pointrange(aes(x = Times, y = Intercept, ymin = Intercept - 1.96*SE, ymax = Intercept + 1.96*SE), position = position_dodge(width = 0.05),size = 0.5, lwd = 0.5,  show.legend = TRUE) + #add points
  
  annotate(geom="point",x = axis.breaks[significant.ps], y = rep(-0.029,length(significant.ps)), colour = "red", alpha = 1.0,size=0.75) + #FOR TANGO AND WALTZ FIGURES
  annotate(geom="point",x = axis.breaks[insignificant.ps], y = rep(-0.029,length(insignificant.ps)), colour = "black", alpha = 1.0,size=0.75) +
  
  
  annotate("text", x = axis.breaks[significant.ps] + c(0,0,-0.075,0,0,0,0,0.05,0,-0.05,0,0,0,-0.075,0,0), y = top.positions[significant.ps]+c(0.015,0.015,0.010,0.020,0.01,0.02,rep(0.005,10)), label = combined.labels[significant.ps],fontface="bold", angle = 0, hjust = 0.5, vjust = 0.75, size = 2) + #FOR TANGO FIGURES
  
  # annotate("text", x = axis.breaks[significant.ps] + c(0,-0.05,-0.05,0.05,0.05,-0.05), y = top.positions[significant.ps] + c(0.01,0.01,0.005,0.01,0,0), label = combined.labels[significant.ps], fontface="bold", angle = 0, hjust = 0.5, vjust = 0.75, size = 2) + #FOR WALTZ FIGURES

  scale_x_continuous(labels=label.vector, breaks = unique(combined.differences.df$Times)) + #change x-axis names of phylostrata to taxonomic split names
  scale_shape_manual(name = "", values = c(22,23)) + #manually set shapes to hollow types
  scale_colour_manual(name = "", values = c(colVermillion, colSkyBlue)) + #manually set line colors
  scale_fill_manual(name = "", values = c(colVermillion, colSkyBlue), guide = guide_legend(ncol = 1)) +
  #guides(nrow = 2) +
  ylim(c(-0.0291,0.155)) +
  ylab("Paired difference in aggregation") + #set y-axis label
  theme(
        panel.grid.minor.x = element_blank(),
        axis.title.x = element_blank(),
        axis.text.y = element_text(size=9, colour="black"),
        axis.title.y = element_text(size=11, colour="black"),
        axis.ticks = element_blank(),
        
        axis.text.x=element_text(angle = 90, size=c(rep(9,7),8,rep(9,14)), hjust = 1, vjust = 0.5,  colour=color.vector), #FOR TANGO AND WALTZ
        #axis.text.x=element_text(angle = 90, size=c(rep(9,7),8,rep(9,14)), hjust = 1, vjust = 0.5, colour="black"),
        #axis.text.x=element_text(angle = 90, size=c(rep(9,7),8,rep(9,13)), hjust = 1, vjust = 0.5, colour=c(rep("black",10),"red","red",rep("black",7))), #FOR WALTZ 
        
        legend.title = element_text(size=9),
        legend.text = element_text(size=9),
        legend.background = element_rect(fill="transparent"),
        legend.position = c(0.41,0.9),
        legend.direction = "vertical",
        #panel.grid.major = element_blank(),
        plot.margin=unit(c(0.1,0.1,0.1,0.1), "in")) #plot specifics for sizes, positions, orientations, etc.
plot.APR.difference

tiff(file = "PairedDifference_TangoAANumInAPR.tif", width =20, height = 10, units = "cm", res = 300) #output plot to tif
plot.APR.difference
dev.off()

# tiff(file = "PairedDifference_WaltzAANumInAPR.tif", width =20, height = 10, units = "cm", res = 300) #output plot to tif
# plot.APR.difference
# dev.off()



```
