---
title: "Amino acid composition comparison summary"
author: "Ben Wilson"
output: html_document
---
  
Load mySQL table info for genes
```{r}
gene.data <- read.table("InputFinal/EnsemblV73_MouseGenes.tab", sep="\t",header=T) 
protein.data <- read.table("InputFinal/EnsemblV73_MouseProteins.tab", sep="\t",header=T)

hydrophobicity.data <- read.table("InputFinal/EnsemblV73_MouseProteins_FILVMW_AAPercent.tab", sep="\t",header=T)
scrambled.data <- read.table("InputFinal/FILVMW_2018ScrambledControls.tab",sep="\t",header=TRUE)

APR.data <- merge(gene.data, protein.data, by.x="MouseGenesPrimaryKey", by.y="MouseGenesForeignKey")
APR.data <- merge(APR.data, hydrophobicity.data[,c("MouseProteinsPrimaryKey","FILVMW_HydrophobicPercentageUsingDispersionAA")], by = "MouseProteinsPrimaryKey")
remove(gene.data)
remove(protein.data)
```

Set phylostrata for intergenic controls
```{r}
APR.data[which(APR.data$GeneOrControlDesignation == "IntergenicNucleotideSeqControl"),]$PS21_GeneFamilyPhylostratum <- "Intergenic" #Give intergenic sequence a categorical phylostratum for comparison
APR.data[which(APR.data$GeneOrControlDesignation == "MaskedIntergenicNucleotideSeqControl"),]$PS21_GeneFamilyPhylostratum <- "Repeat-masked" #Give repeat-masked intergenic sequence a categorical phylostratum
```

```{r}
APR.data[which(APR.data$EnsemlGeneUID=="ENSMUSG00000067543"),]$PS21_GeneFamilyNumber <- APR.data[which(APR.data$EnsemlGeneUID=="ENSMUSG00000052262"),]$PS21_GeneFamilyNumber #Fix gene family for two pahari genes

```

Factorize gene family number and phylostratum
```{r}
APR.data$PS21_GeneFamilyNumber <- as.factor(APR.data$PS21_GeneFamilyNumber) #factorize gene family ID to prevent numeric interpretation by R
APR.data$PS21_GeneFamilyPhylostratum <- as.factor(APR.data$PS21_GeneFamilyPhylostratum) #factorize variable to ensure that it is not interpreted as numeric variable
APR.data$PS21_GeneFamilyPhylostratum <- with(APR.data, factor(PS21_GeneFamilyPhylostratum, levels = c(sort(unique(APR.data$PS21_GeneFamilyPhylostratum)),"Intergenic","Repeat-masked"))) # set levels of categorical phylostratum variable
APR.data$PS21_ExcludedGeneBinary <- as.factor(APR.data$PS21_ExcludedGeneBinary) #factorize binary category for gene families with ambiguous age

```


Remove phylostratum 0 (unclassifiable), gene families with ambiguous age, genes without rat homologs, and ORFan proteins
```{r}
APR.data <- APR.data[which(APR.data$PS21_GeneFamilyPhylostratum != "0" & APR.data$PS21_GeneFamilyPhylostratum != "21" & APR.data$PS21_ExcludedGeneBinary != "1" ),] #exclude unclassifiable genes, false-positive rich ORFans
paralogs.list <- c("ENSMUSG00000073631","ENSMUSG00000079800","ENSMUSG00000094728","ENSMUSG00000095240","ENSMUSG00000095366","ENSMUSG00000095450","ENSMUSG00000095787","ENSMUSG00000096100","ENSMUSG00000096850")
APR.data <- APR.data[which(!(APR.data$EnsemblGeneUID %in% paralogs.list)),]#remove genes with known paralogy to other genes
```

Trim data frame to only the information needed
```{r}
genes <- APR.data[which(APR.data$GeneOrControlDesignation == "CodingGene"),c("EnsemblGeneUID","PS21_GeneFamilyNumber","PS21_GeneFamilyPhylostratum","TangoPentamerAAInAPRsDensity","MannigeHydrophobicityPercent","BoussauThermophilicPercent","FILVMW_HydrophobicPercentageUsingDispersionAA")] #subset out genes
scrambled <- APR.data[which(APR.data$GeneOrControlDesignation == "ScrambledByAminoAcidControl"),c("EnsemblGeneUID","PS21_GeneFamilyNumber","PS21_GeneFamilyPhylostratum","TangoPentamerAAInAPRsDensity","MannigeHydrophobicityPercent","BoussauThermophilicPercent","FILVMW_HydrophobicPercentageUsingDispersionAA")] #subset out scrambled controls
intergenic <- APR.data[which(APR.data$GeneOrControlDesignation == "IntergenicNucleotideSeqControl"),c("EnsemblGeneUID","PS21_GeneFamilyNumber","PS21_GeneFamilyPhylostratum","TangoPentamerAAInAPRsDensity","MannigeHydrophobicityPercent","BoussauThermophilicPercent","FILVMW_HydrophobicPercentageUsingDispersionAA")] #subset out intergenic
masked <- APR.data[which(APR.data$GeneOrControlDesignation == "MaskedIntergenicNucleotideSeqControl"),c("EnsemblGeneUID","PS21_GeneFamilyNumber","PS21_GeneFamilyPhylostratum","TangoPentamerAAInAPRsDensity","MannigeHydrophobicityPercent","BoussauThermophilicPercent","FILVMW_HydrophobicPercentageUsingDispersionAA")] #subset out intergenic

scrambled.mean <- aggregate(scrambled.data$TangoPentamerAAInAPRsDensity, by = list(scrambled.data$EnsemblGeneUID), FUN = mean, data=scrambled.data)
colnames(scrambled.mean) <- c("EnsemblGeneUID","ScrambledMeanAPR")

scrambled <- merge(scrambled, scrambled.mean, by = "EnsemblGeneUID")

remove(APR.data) #clear large data frame with unnecessary information
```

Use two-parameter Box-Cox transform to reduce skewness
```{r}
library(geoR)

bcFit.TangoNumAA <- boxcoxfit(genes$TangoPentamerAAInAPRsDensity, lambda2 =  NULL) #Fit Box-Cox transform to each aggregation metric
lambda1.TangoNumAA <- unname(bcFit.TangoNumAA$lambda[1])
lambda1.TangoNumAA

bcFit.Oiliness <- boxcoxfit(genes$MannigeHydrophobicityPercent, lambda2 =  NULL) #Fit Box-Cox transform to each metric
lambda1.Oiliness <- unname(bcFit.Oiliness$lambda[1])
lambda1.Oiliness

bcFit.Thermo <- boxcoxfit(genes$BoussauThermophilicPercent, lambda2 =  NULL) #Fit Box-Cox transform to each metric
lambda1.Thermo <- unname(bcFit.Thermo$lambda[1])
lambda1.Thermo

bcFit.Hydro <- boxcoxfit(genes$FILVMW_HydrophobicPercentageUsingDispersionAA, lambda2 =  NULL) #Fit Box-Cox transform to each metric
lambda1.Hydro <- unname(bcFit.Hydro$lambda[1])
lambda1.Hydro


bc.transform <- function(x,l1,l2){ #Create function to transform data according to fit parameters
  x.transform <- ((x+l2)^l1 - 1)/l1
  return(x.transform)
}


scrambled$bcTangoNumAA <- bc.transform(scrambled$ScrambledMeanAPR, lambda1.TangoNumAA, 0) #tranform metrics with fitted parameters
intergenic$bcTangoNumAA <- bc.transform(intergenic$TangoPentamerAAInAPRsDensity, lambda1.TangoNumAA, 0) #...
masked$bcTangoNumAA <- bc.transform(masked$TangoPentamerAAInAPRsDensity, lambda1.TangoNumAA, 0) #...

genes$bcOiliness <- bc.transform(genes$MannigeHydrophobicityPercent, lambda1.Oiliness, 0) #...
intergenic$bcOiliness <- bc.transform(intergenic$MannigeHydrophobicityPercent, lambda1.Oiliness, 0) #...
masked$bcOiliness <- bc.transform(masked$MannigeHydrophobicityPercent, lambda1.Oiliness, 0) #...


genes$bcThermo <- bc.transform(genes$BoussauThermophilicPercent, lambda1.Thermo, 0) #...
intergenic$bcThermo <- bc.transform(intergenic$BoussauThermophilicPercent, lambda1.Thermo, 0) #...
masked$bcThermo <- bc.transform(masked$BoussauThermophilicPercent, lambda1.Thermo, 0) #...

genes$bcHydro <- bc.transform(genes$FILVMW_HydrophobicPercentageUsingDispersionAA, lambda1.Hydro, 0) #...
intergenic$bcHydro <- bc.transform(intergenic$FILVMW_HydrophobicPercentageUsingDispersionAA, lambda1.Hydro, 0) #...
masked$bcHydro <- bc.transform(masked$FILVMW_HydrophobicPercentageUsingDispersionAA, lambda1.Hydro, 0) #...
```

Generate family of linear models for each phylostratum controlling for gene family variance
```{r}

#**********NLME FITS*********
library(nlme)

#Tango scrambled fits
fit.APR.scrambled <- lme(bcTangoNumAA ~ PS21_GeneFamilyPhylostratum, random = ~ 1|PS21_GeneFamilyNumber, data = scrambled, control = list(opt = "optim"), na.action = na.exclude) #fit linear mixed-effects model for genes with gene family as random error term
summary(fit.APR.scrambled) 

fit.APR.intergenic <- lm(bcTangoNumAA ~ 1 , data = intergenic) #fit standard linear model to intergenic controls
fit.APR.masked <- lm(bcTangoNumAA ~ 1 , data = masked) #...same for repeat-masked controls



#Hydrophobicity fits
fit.oiliness.genes <- lme(bcOiliness ~ PS21_GeneFamilyPhylostratum , random = ~ 1|PS21_GeneFamilyNumber, data = genes, control = list(opt = "optim"), na.action = na.exclude)
summary(fit.oiliness.genes) #fit a mixed-effects model with gene family as a random effect term

fit.oiliness.intergenic <- lm(bcOiliness ~ 1, data = intergenic)
fit.oiliness.masked <- lm(bcOiliness ~ 1, data = masked)


#Thermophilicity fits
fit.thermo.genes <- lme(bcThermo ~ PS21_GeneFamilyPhylostratum , random = ~ 1|PS21_GeneFamilyNumber, data = genes, control = list(opt = "optim"), na.action = na.exclude)
summary(fit.thermo.genes) #fit a mixed-effects model with gene family as a random effect term

fit.thermo.intergenic <- lm(bcThermo ~ 1, data = intergenic)
fit.thermo.masked <- lm(bcThermo ~ 1, data = masked)

#Hydrophobicity fits
fit.hydro.genes <- lme(bcHydro ~ PS21_GeneFamilyPhylostratum , random = ~ 1|PS21_GeneFamilyNumber, data = genes, control = list(opt = "optim"), na.action = na.exclude)
summary(fit.hydro.genes) #fit a mixed-effects model with gene family as a random effect term

fit.hydro.intergenic <- lm(bcHydro ~ 1, data = intergenic)
fit.hydro.masked <- lm(bcHydro ~ 1, data = masked)

```

Build data frames from linear model outptut
```{r}
# Put model estimates into data frames for aggregation data
last.index.scrambled <- length(fixed.effects(fit.APR.scrambled))  #store number of PS that have been fit for scrambled
scrambled.df.APR <- data.frame(Phylostratum = names(fixed.effects(fit.APR.scrambled)[1:last.index.scrambled]), Intercept = unname(fixed.effects(fit.APR.scrambled)[1:last.index.scrambled]), SE = unname(summary(fit.APR.scrambled)$tTable[,2][1:last.index.scrambled]), Model = "Scrambled controls", Metric = "Aggregation") #create data frame to store lm results for the scrambled fit
remove(fit.APR.scrambled)

intergenic.df.APR <- data.frame(Phylostratum = c("Intergenic"), Intercept = summary(fit.APR.intergenic)$coefficients[1], SE = summary(fit.APR.intergenic)$coefficients[2], Model = "Intergenic", Metric = "Aggregation") #create data frame to store intergenic fit
remove(fit.APR.intergenic)

masked.df.APR <- data.frame(Phylostratum = c("Repeat-masked"), Intercept = summary(fit.APR.masked)$coefficients[1], SE = summary(fit.APR.masked)$coefficients[2], Model = "Repeat-masked", Metric = "Aggregation") #create data frame to store repeat-masked intergenic fit
remove(fit.APR.masked)


# Put model estimates into data frames for hydrophobicity data
last.index.genes.oiliness <- length(fixed.effects(fit.oiliness.genes))  #store number of PS that have been fit for genes
genes.df.oiliness <- data.frame(Phylostratum = names(fixed.effects(fit.oiliness.genes)[1:last.index.genes.oiliness]), Intercept = unname(fixed.effects(fit.oiliness.genes)[1:last.index.genes.oiliness]), SE = unname(summary(fit.oiliness.genes)$tTable[,2][1:last.index.genes.oiliness]), Model = "Genes", Metric = "Oiliness") #create data frame to store lm results for the genes fit
remove(fit.oiliness.genes)

intergenic.df.oiliness <- data.frame(Phylostratum = c("Intergenic"), Intercept = summary(fit.oiliness.intergenic)$coefficients[1], SE = summary(fit.oiliness.intergenic)$coefficients[2], Model = "Intergenic", Metric = "Oiliness") #create data frame to store intergenic fit
remove(fit.oiliness.intergenic)

masked.df.oiliness <- data.frame(Phylostratum = c("Repeat-masked"), Intercept = summary(fit.oiliness.masked)$coefficients[1], SE = summary(fit.oiliness.masked)$coefficients[2], Model = "Repeat-masked", Metric = "Oiliness") #create data frame to store repeat-masked intergenic fit
remove(fit.oiliness.masked)


# Put model estimates into data frames for thermophilicity data
last.index.genes.thermo <- length(fixed.effects(fit.thermo.genes))  #store number of PS that have been fit for genes
genes.df.thermo <- data.frame(Phylostratum = names(fixed.effects(fit.thermo.genes)[1:last.index.genes.thermo]), Intercept = unname(fixed.effects(fit.thermo.genes)[1:last.index.genes.thermo]), SE = unname(summary(fit.thermo.genes)$tTable[,2][1:last.index.genes.thermo]), Model = "Genes", Metric = "Thermophilicity") #create data frame to store lm results for the genes fit
remove(fit.thermo.genes)

intergenic.df.thermo <- data.frame(Phylostratum = c("Intergenic"), Intercept = summary(fit.thermo.intergenic)$coefficients[1], SE = summary(fit.thermo.intergenic)$coefficients[2], Model = "Intergenic", Metric = "Thermophilicity") #create data frame to store intergenic fit
remove(fit.thermo.intergenic
       )
masked.df.thermo <- data.frame(Phylostratum = c("Repeat-masked"), Intercept = summary(fit.thermo.masked)$coefficients[1], SE = summary(fit.thermo.masked)$coefficients[2], Model = "Repeat-masked", Metric = "Thermophilicity") #create data frame to store repeat-masked intergenic fit
remove(fit.thermo.masked)


# Put model estimates into data frames for hydrophobicity data
last.index.genes.hydro <- length(fixed.effects(fit.hydro.genes))  #store number of PS that have been fit for genes
genes.df.hydro <- data.frame(Phylostratum = names(fixed.effects(fit.hydro.genes)[1:last.index.genes.hydro]), Intercept = unname(fixed.effects(fit.hydro.genes)[1:last.index.genes.hydro]), SE = unname(summary(fit.hydro.genes)$tTable[,2][1:last.index.genes.hydro]), Model = "Genes", Metric = "Hydrophobicity") #create data frame to store lm results for the genes fit
remove(fit.hydro.genes)

intergenic.df.hydro <- data.frame(Phylostratum = c("Intergenic"), Intercept = summary(fit.hydro.intergenic)$coefficients[1], SE = summary(fit.hydro.intergenic)$coefficients[2], Model = "Intergenic", Metric = "Hydrophobicity") #create data frame to store intergenic fit
remove(fit.hydro.intergenic
       )
masked.df.hydro <- data.frame(Phylostratum = c("Repeat-masked"), Intercept = summary(fit.hydro.masked)$coefficients[1], SE = summary(fit.hydro.masked)$coefficients[2], Model = "Repeat-masked", Metric = "Hydrophobicity") #create data frame to store repeat-masked intergenic fit
remove(fit.hydro.masked)



```

Adjust means for each phylostrata by incrementing by the first mean (adjust for R convention), combine data frames per categorization
```{r}
# Adjust intecepts, combine data frames
scrambled.df.APR$Intercept[2:last.index.scrambled] <- scrambled.df.APR$Intercept[2:last.index.scrambled] + scrambled.df.APR$Intercept[1] #increment means by first means to correct for R convention for storing fit results

genes.df.oiliness$Intercept[2:last.index.genes.oiliness] <- genes.df.oiliness$Intercept[2:last.index.genes.oiliness] + genes.df.oiliness$Intercept[1] #...same for other fits

genes.df.thermo$Intercept[2:last.index.genes.thermo] <- genes.df.thermo$Intercept[2:last.index.genes.thermo] + genes.df.thermo$Intercept[1] #...same for other fits

genes.df.hydro$Intercept[2:last.index.genes.hydro] <- genes.df.hydro$Intercept[2:last.index.genes.hydro] + genes.df.hydro$Intercept[1] #...same for other fits


#Define reverse transform
bc.reverse.transform <- function(x,l1,l2){
  x.reverse.transform <- (x*l1+1)^(1/l1)-l2
  return(x.reverse.transform)
}

#Reverse Box-Cox transform so data are in human-readable format

scrambled.df.APR$Upper <- bc.reverse.transform((scrambled.df.APR$Intercept + scrambled.df.APR$SE), lambda1.TangoNumAA, 0)
scrambled.df.APR$Lower <- bc.reverse.transform((scrambled.df.APR$Intercept - scrambled.df.APR$SE), lambda1.TangoNumAA, 0)
scrambled.df.APR$Intercept <- bc.reverse.transform(scrambled.df.APR$Intercept, lambda1.TangoNumAA, 0)
intergenic.df.APR$Upper <- bc.reverse.transform((intergenic.df.APR$Intercept + intergenic.df.APR$SE), lambda1.TangoNumAA, 0)
intergenic.df.APR$Lower <- bc.reverse.transform((intergenic.df.APR$Intercept - intergenic.df.APR$SE), lambda1.TangoNumAA, 0)
intergenic.df.APR$Intercept <- bc.reverse.transform(intergenic.df.APR$Intercept, lambda1.TangoNumAA, 0)
masked.df.APR$Upper <- bc.reverse.transform((masked.df.APR$Intercept + masked.df.APR$SE), lambda1.TangoNumAA, 0)
masked.df.APR$Lower <- bc.reverse.transform((masked.df.APR$Intercept - masked.df.APR$SE), lambda1.TangoNumAA, 0)
masked.df.APR$Intercept <- bc.reverse.transform(masked.df.APR$Intercept, lambda1.TangoNumAA, 0)

genes.df.oiliness$Upper <- bc.reverse.transform((genes.df.oiliness$Intercept + genes.df.oiliness$SE), lambda1.Oiliness, 0)
genes.df.oiliness$Lower <- bc.reverse.transform((genes.df.oiliness$Intercept - genes.df.oiliness$SE), lambda1.Oiliness, 0)
genes.df.oiliness$Intercept <- bc.reverse.transform(genes.df.oiliness$Intercept, lambda1.Oiliness, 0)
intergenic.df.oiliness$Upper <- bc.reverse.transform((intergenic.df.oiliness$Intercept + intergenic.df.oiliness$SE), lambda1.Oiliness, 0)
intergenic.df.oiliness$Lower <- bc.reverse.transform((intergenic.df.oiliness$Intercept - intergenic.df.oiliness$SE), lambda1.Oiliness, 0)
intergenic.df.oiliness$Intercept <- bc.reverse.transform(intergenic.df.oiliness$Intercept, lambda1.Oiliness, 0)
masked.df.oiliness$Upper <- bc.reverse.transform((masked.df.oiliness$Intercept + masked.df.oiliness$SE), lambda1.Oiliness, 0)
masked.df.oiliness$Lower <- bc.reverse.transform((masked.df.oiliness$Intercept - masked.df.oiliness$SE), lambda1.Oiliness, 0)
masked.df.oiliness$Intercept <- bc.reverse.transform(masked.df.oiliness$Intercept, lambda1.Oiliness, 0)


genes.df.thermo$Upper <- bc.reverse.transform((genes.df.thermo$Intercept + genes.df.thermo$SE), lambda1.Thermo, 0)
genes.df.thermo$Lower <- bc.reverse.transform((genes.df.thermo$Intercept - genes.df.thermo$SE), lambda1.Thermo, 0)
genes.df.thermo$Intercept <- bc.reverse.transform(genes.df.thermo$Intercept, lambda1.Thermo, 0)
intergenic.df.thermo$Upper <- bc.reverse.transform((intergenic.df.thermo$Intercept + intergenic.df.thermo$SE), lambda1.Thermo, 0)
intergenic.df.thermo$Lower <- bc.reverse.transform((intergenic.df.thermo$Intercept - intergenic.df.thermo$SE), lambda1.Thermo, 0)
intergenic.df.thermo$Intercept <- bc.reverse.transform(intergenic.df.thermo$Intercept, lambda1.Thermo, 0)
masked.df.thermo$Upper <- bc.reverse.transform((masked.df.thermo$Intercept + masked.df.thermo$SE), lambda1.Thermo, 0)
masked.df.thermo$Lower <- bc.reverse.transform((masked.df.thermo$Intercept - masked.df.thermo$SE), lambda1.Thermo, 0)
masked.df.thermo$Intercept <- bc.reverse.transform(masked.df.thermo$Intercept, lambda1.Thermo, 0)

genes.df.hydro$Upper <- bc.reverse.transform((genes.df.hydro$Intercept + genes.df.hydro$SE), lambda1.Hydro, 0)
genes.df.hydro$Lower <- bc.reverse.transform((genes.df.hydro$Intercept - genes.df.hydro$SE), lambda1.Hydro, 0)
genes.df.hydro$Intercept <- bc.reverse.transform(genes.df.hydro$Intercept, lambda1.Hydro, 0)
intergenic.df.hydro$Upper <- bc.reverse.transform((intergenic.df.hydro$Intercept + intergenic.df.hydro$SE), lambda1.Hydro, 0)
intergenic.df.hydro$Lower <- bc.reverse.transform((intergenic.df.hydro$Intercept - intergenic.df.hydro$SE), lambda1.Hydro, 0)
intergenic.df.hydro$Intercept <- bc.reverse.transform(intergenic.df.hydro$Intercept, lambda1.Hydro, 0)
masked.df.hydro$Upper <- bc.reverse.transform((masked.df.hydro$Intercept + masked.df.hydro$SE), lambda1.Hydro, 0)
masked.df.hydro$Lower <- bc.reverse.transform((masked.df.hydro$Intercept - masked.df.hydro$SE), lambda1.Hydro, 0)
masked.df.hydro$Intercept <- bc.reverse.transform(masked.df.hydro$Intercept, lambda1.Hydro, 0)


combined.df <- rbind(scrambled.df.APR, genes.df.oiliness, genes.df.thermo, genes.df.hydro, intergenic.df.APR, intergenic.df.oiliness, intergenic.df.thermo, intergenic.df.hydro, masked.df.APR, masked.df.oiliness, masked.df.thermo, masked.df.hydro) #combine result data frames
combined.df$Phylostratum <- replace(as.character(combined.df$Phylostratum),1:(last.index.scrambled+last.index.genes.oiliness+last.index.genes.thermo+last.index.genes.hydro),c(1:last.index.scrambled,1:last.index.genes.oiliness,1:last.index.genes.thermo,1:last.index.genes.hydro)) #change PS names for the scrambled and gene fits to numeric characters
combined.df$Phylostratum <- with(combined.df, factor(Phylostratum, levels = c(1:max(last.index.genes.thermo,last.index.genes.oiliness,last.index.scrambled,last.index.genes.hydro),"Intergenic","Repeat-masked"))) #change factor levels for PS for combined results data frame

```


Make plots to visualize effects of different categorization
```{r}
# Plots for each categorization
library("grid")
library("ggplot2")


#Create colorblind friendly palette
colorblindPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
colBlack <- colorblindPalette[1]
colOrange <- colorblindPalette[2]
colSkyBlue <- colorblindPalette[3]
colGreen <- colorblindPalette[4]
colYellow <- colorblindPalette[5]
colDarkBlue <- colorblindPalette[6]
colVermillion <- colorblindPalette[7]
colPink <- colorblindPalette[8]

interval.times <- read.table("split_interval_times.txt")
axis.breaks <- -log10(interval.times$V1)
linear.scale <- c(-1*interval.times$V1)

log.scale <- c(axis.breaks,-6.9,-6.85)
label.vector <- c("Cellular org.","Eukaryota","Opisthokonta","Holozoa","Metazoa","","","","","","Vertebrata","Euteleostomi","Tetrapoda","Amniota","Mammalia","Eutheria","Boreoeutheria\n","Euarchontoglires","Rodentia","Mus","Intergenic","Masked")

combined.df$Times <-c(rep(log.scale[1:20],4),rep(-6.9,4),rep(-6.85,4))


thermophilicity.plot <- ggplot(combined.df[which(combined.df$Metric=="Thermophilicity"),], aes(colour = Metric, fill = Metric, shape = Model)) #Create ggplot object
thermophilicity.plot <- thermophilicity.plot + 
  geom_pointrange(aes(x = Times, y = Intercept, ymin = Lower, ymax = Upper), size = 0.5, lwd = 1,  show.legend=FALSE) + #add point between error bars
  geom_vline(xintercept = -7, color = "darkgrey", linetype = 2) +
  scale_x_continuous(labels=label.vector, breaks = unique(combined.df$Times)) + #change x-axis names of phylostrata to taxonomic split names
  scale_shape_manual(name = "", values = c(21,23,24)) + #manually set hollow shapes
  scale_colour_manual(name = "", values = c(colPink), guide = FALSE) + #manually set line colors
  scale_fill_manual(name = "", values = c(colPink), guide = FALSE) + #manually set fill colors
  ylim(c(0.3,0.45))+
  ylab("Thermophilicity\n")+
  theme( 
    panel.grid.minor.x = element_blank(),
    axis.title.x = element_blank(), 
    axis.text.y = element_text(size=10),
    axis.title.y = element_text(size=11),
    axis.ticks = element_blank(),
    axis.text.x=element_text(angle = 90, size=c(rep(9,7),8,rep(9,14)), hjust = 1, vjust = 0.5, colour=c(rep("black",7),"lightslategrey",rep("black",14))),
    legend.title = element_text(size=11),
    legend.text = element_text(size=8),
    legend.background = element_rect(fill="transparent"),
    legend.position = c(0.5,0.95),
    legend.direction = "horizontal",
    plot.margin=unit(c(0.1,0.1,0.1,0.1), "cm")) #plot specifics for sizes, positions, orientations, etc.
thermophilicity.plot



oiliness.plot <- ggplot(combined.df[which(combined.df$Metric=="Oiliness"),], aes(colour = Metric, fill = Metric, shape = Model)) #Create ggplot object
oiliness.plot <- oiliness.plot + 
  geom_pointrange(aes(x = Times, y = Intercept, ymin = Lower, ymax = Upper), size = 0.5, lwd = 1,  show.legend=FALSE) + #add point between error bars
  geom_vline(xintercept = -7, color = "darkgrey", linetype = 2) +
  scale_x_continuous(labels=label.vector, breaks = unique(combined.df$Times)) + #change x-axis names of phylostrata to taxonomic split names
  scale_shape_manual(name = "", values = c(21,23,24)) + #manually set hollow shapes
  scale_colour_manual(name = "", values = c(colBlack), guide = FALSE) + #manually set line colors
  scale_fill_manual(name = "", values = c(colBlack), guide = FALSE) + #manually set fill colors
  ylim(c(0.15,0.3))+
  ylab("Oiliness %FILV\n") +
  theme( 
    panel.grid.minor.x = element_blank(),
    axis.title.x = element_blank(), 
    axis.text.y = element_text(size=10),
    axis.title.y = element_text(size=11),
    axis.ticks = element_blank(),
    axis.text.x=element_blank(),
    legend.title = element_text(size=11),
    legend.text = element_text(size=8),
    legend.background = element_rect(fill="transparent"),
    legend.position = c(0.5,0.95),
    legend.direction = "horizontal",
    plot.margin=unit(c(0.1,0.1,0.1,0.1), "cm")) #plot specifics for sizes, positions, orientations, etc.
oiliness.plot

hydrophobicity.plot <- ggplot(combined.df[which(combined.df$Metric=="Hydrophobicity"),], aes(colour = Metric, fill = Metric, shape = Model)) #Create ggplot object
hydrophobicity.plot <- hydrophobicity.plot + 
  geom_pointrange(aes(x = Times, y = Intercept, ymin = Lower, ymax = Upper), size = 0.5, lwd = 1,  show.legend=FALSE) + #add point between error bars
  geom_vline(xintercept = -7, color = "darkgrey", linetype = 2) +
  scale_x_continuous(labels=label.vector, breaks = unique(combined.df$Times)) + #change x-axis names of phylostrata to taxonomic split names
  scale_shape_manual(name = "", values = c(21,23,24)) + #manually set hollow shapes
  scale_colour_manual(name = "", values = c(colGreen), guide = FALSE) + #manually set line colors
  scale_fill_manual(name = "", values = c(colGreen), guide = FALSE) + #manually set fill colors
  #ylim(c(0.15,0.3))+
  ylab("Hydrophobicity\n") +
  theme( 
    panel.grid.minor.x = element_blank(),
    axis.title.x = element_blank(), 
    axis.text.y = element_text(size=10),
    axis.title.y = element_text(size=11),
    axis.ticks = element_blank(),
    axis.text.x=element_blank(),
    legend.title = element_text(size=11),
    legend.text = element_text(size=8),
    legend.background = element_rect(fill="transparent"),
    legend.position = c(0.5,0.95),
    legend.direction = "horizontal",
    plot.margin=unit(c(0.1,0.1,0.1,0.1), "cm")) #plot specifics for sizes, positions, orientations, etc.
hydrophobicity.plot



aggregation.plot <- ggplot(combined.df[which(combined.df$Metric=="Aggregation"),], aes(colour = Metric, fill = Metric, shape = Model)) #Create ggplot object
aggregation.plot <- aggregation.plot + 
  geom_pointrange(aes(x = Times, y = Intercept, ymin = Lower, ymax = Upper), size = 0.5, lwd = 1,  show.legend=FALSE) + #add point between error bars
  geom_vline(xintercept = -7, color = "darkgrey", linetype = 2) +
  scale_x_continuous(labels=label.vector, breaks = unique(combined.df$Times)) + #change x-axis names of phylostrata to taxonomic split names
  scale_shape_manual(name = "", values = c(22,23,24)) + #manually set hollow shapes
  scale_colour_manual(name = "", values = c(colVermillion), guide = FALSE) + #manually set line colors
  scale_fill_manual(name = "", values = c(colVermillion), guide = FALSE) + #manually set fill colors
  ylim(c(0.0,0.2))+
  ylab("Aggregation\n") +
  theme( 
    panel.grid.minor.x = element_blank(),
    axis.title.x = element_blank(), 
    axis.text.y = element_text(size=10),
    axis.title.y = element_text(size=11),
    axis.ticks = element_blank(),
    axis.text.x=element_text(angle = 90, size=c(rep(9,7),8,rep(9,14)), hjust = 1, vjust = 0.5, colour=c(rep("black",7),"lightslategrey",rep("black",12),colDarkBlue,colSkyBlue)),
    legend.title = element_text(size=11),
    legend.text = element_text(size=8),
    legend.background = element_rect(fill="transparent"),
    legend.position = c(0.5,0.95),
    legend.direction = "horizontal",
    plot.margin=unit(c(0.1,0.1,0.1,0.1), "cm")) #plot specifics for sizes, positions, orientations, etc.
aggregation.plot

load("ISD_plot.grob") #easier to run ISD script to generate plot.ISD, then grab from memory, as below


tiff(file = "AA_Composition_Summary.tif", width =20, height = 20, units = "cm", res = 300) #output plot to tif
grid.newpage()
grid.draw(rbind(ggplotGrob(aggregation.plot + theme(axis.text.x=element_blank())),ggplotGrob(hydrophobicity.plot), ggplotGrob(plot.ISD+theme(axis.text.x=element_blank())),ggplotGrob(thermophilicity.plot),size = "last"))
dev.off()


```